\documentclass[11pt,a4paper]{article}

\usepackage{amsmath,amssymb,amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{doi}
\geometry{margin=2.5cm}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}

\newcommand{\bstar}{\beta^{*}}
\newcommand{\ImH}{\mathrm{Im}\,\mathbb{H}}
\newcommand{\su}{\mathfrak{su}}
\newcommand{\SU}{\mathrm{SU}}
\newcommand{\SO}{\mathrm{SO}}
\newcommand{\Heis}{\mathrm{Heis}_{3}(\mathbb{Z}/q\mathbb{Z})}
\newcommand{\sigc}{\Sigma_{c}}
\newcommand{\dpair}{\delta_{\mathrm{pair}}}

\title{Three Stable Directions from Quaternionic Minimality:\\
Derivation of $\Sigma_{c}(n_{3}) = 3$ from Born--Infeld Fibre Admissibility\\
\large O23 --- Spectral Admissibility Sub-programme}

\author{J\'{e}r\^{o}me Beau\\
\small Independent Researcher}

\date{2026}

\begin{document}

  \maketitle

  \begin{abstract}
    O21~\cite{Beau2026a25} defined the observable shell $n_{3}$ via the threshold condition
    $\Sigma_{c}(n_{3}) = 3$ and used it as structural input for the fibre-level transfer
    $\dpair \to \bstar$.
    O22 established that Born--Infeld projection locking forces saturation to occur on a
    BFS shell, closing the question ``why a shell?''
    The present paper closes the complementary question left open by O21: why is the
    relevant saturation threshold three-dimensional, i.e.\ why $\Sigma_{c}(n_{3}) = 3$
    rather than $2$ or $4$?
    We show that this value is not an arithmetic input but a structural consequence of
    Born--Infeld fibre admissibility: the value three is not a spectral artefact but
    the unique dimension of the imaginary subspace of the minimal admissible associative
    non-commutative algebra.
    The argument proceeds in three steps.
    First, an exclusion lemma rules out abelian realizations of the admissible neutral
    sector and limits the candidate structures to finite subgroups of $\SU(2)$.
    Second, a quaternionic minimality theorem --- anchored in the Hurwitz-to-$\su(2)$
    bridge established in SpectralCapacity~\cite{Beau2026a2} --- identifies
    $\mathbb{H}$ as the minimal associative non-commutative structure compatible with
    Born--Infeld parity, and shows that the admissible neutral directions correspond to
    the imaginary subspace $\ImH$, which carries exactly three independent axes.
    Third, a spectral realization section shows that the ADE binary graphs physically
    relevant to the programme realize this abstract ``$3$'' as exactly three distinct
    non-trivial spectral levels, confirming the structural derivation at the observable
    level.
    The result is conditional on the absence of additional effective symmetries beyond
    Born--Infeld parity, in accordance with the explicit scope of O18~\cite{Beau2026a22}.
  \end{abstract}

  \paragraph{Keywords.}
    spectral admissibility; quaternionic minimality; Born–Infeld constraint;
    SU(2) representations; imaginary quaternions; Weil representation; ADE classification

    \clearpage
    \tableofcontents

    \clearpage
  \section{Position of the Problem after O21--O22}
  \label{sec:position}

  \subsection{What O21 established and what it left open}
    \label{ssec:o21-status}

    O21~\cite{Beau2026a25} established the shell-level saturation mechanism and derived
    the fibre-level transfer $\dpair \to \bstar$.
    A central ingredient was the threshold condition
    \begin{equation}
      \label{eq:threshold}
      \Sigma_{c}(n_{3}) = 3,
    \end{equation}
    which defines the observable shell $n_{3}$ used throughout the cascade analysis.
    O21 used~\eqref{eq:threshold} as structural input: the value $3$ was interpreted as
    the number of physically stable directions selected by $\Pi$ in the Weil realisation
    of $\Heis$, but was not derived from the Born--Infeld structure.
    O21~§6.1 explicitly identified the derivation of~\eqref{eq:threshold} as the open
    problem of O23.

    O22~\cite{Beau2026a26} resolved the complementary question.
    Theorem~4.8 of O22 shows that projection locking under Born--Infeld spectral
    admissibility forces saturation to occur on a BFS shell ($n_{\mathrm{sat}} \in
  \mathbb{N}$), answering ``why a shell?''
    O22~§6.1 confirms that the question ``why $\Sigma_{c}(n_{3}) = 3$?'' thereby becomes
    well-posed, and assigns it to the present paper.
    The present result removes the last remaining free structural integer in the spectral
    admissibility pipeline.

  \subsection{The three distinct ``3''s that must not be conflated}
    \label{ssec:three-threes}

    The number $3$ appears in three structurally distinct places in the programme, and
    O23 must keep them separate throughout.

    \begin{enumerate}
      \item $\dim_{\mathbb{R}} \ImH = 3$: the real dimension of the imaginary quaternions,
      i.e.\ the subspace of $\mathbb{H}$ orthogonal to $\mathbb{R} \cdot \mathbf{1}$ and
      closed under the commutator.
      This is the algebraic ``$3$'' that O23 derives.

      \item $\dim_{\mathbb{R}} \su(2) = 3$: the dimension of the Lie algebra $\su(2)$
      (equivalently, the dimension of $\SU(2)$ as a Lie group).
      This coincides with~(1) because $\su(2) \cong \ImH$ as real Lie algebras, but the
      identification must be derived, not assumed.

      \item $\Sigma_{c}(n_{3}) = 3$: the observable threshold of O21.
      This is the spectral ``$3$'' that the present paper derives as the signature of~(1)
      at the level of the Gram--Schmidt capacity count.
    \end{enumerate}

    The causal chain of O23 is: $(1) \Rightarrow (2) \Rightarrow (3)$.
    Collapsing these three occurrences into a single unexplained ``$3$'' would reproduce
    exactly the gap that O23 is designed to close.

  \subsection{Scope and what this paper does not claim}
    \label{ssec:scope}

    O23 derives the value $3$ in~\eqref{eq:threshold} as a structural consequence of
    Born--Infeld fibre admissibility.
    It does \emph{not} determine the numerical value of $n_{3}$ as a function of
    $(c_{\mathrm{BI}}, \lambda_{n}, \text{graph geometry})$.
    That determination requires the extended numerical programme described in
    O22~§6.2 and remains an open direction.
    The present result is therefore a structural closure, not a quantitative one.

  \section{Exclusion Lemma: No Abelian Realisation of the Admissible Neutral Sector}
  \label{sec:exclusion}

  \subsection{Setup: neutral generators and the spinorial constraint}
    \label{ssec:neutral-setup}

    We recall the setup of the Admissibility Sub-programme~\cite{Beau2026a1,Beau2026a2}.
    Let $G$ be a finite group, $S = S^{-1} \subset G$ a symmetric generating set, and
    $\rho \colon G \to \SU(2)$ an irreducible two-dimensional representation.
    The \emph{neutrality condition} is
    \begin{equation}
      \label{eq:neutrality}
      \chi_{\rho}(s) := \mathrm{Tr}\,\rho(s) = 0 \quad \forall\, s \in S.
    \end{equation}
    This condition is the spectral admissibility constraint under bounded relational flux:
    it selects the generators that remain compatible with the Born--Infeld capacity
    bound~\cite{Beau2026a1,Beau2026a2}.
    By the Pauli tracelessness identity, each $\rho(s)$ satisfying~\eqref{eq:neutrality}
    can be written
    \begin{equation}
      \label{eq:pauli-axis}
      \rho(s) = \mathrm{i}(\vec{u}_{s} \cdot \vec{\sigma}), \quad
      \vec{u}_{s} \in S^{2} \subset \mathbb{R}^{3},
    \end{equation}
    where $\vec{\sigma} = (\sigma_{x}, \sigma_{y}, \sigma_{z})$ are the Pauli matrices.
    Each neutral generator therefore determines a unique axis $\pm \vec{u}_{s} \in
  \mathbb{RP}^{2}$ in the three-sphere~\cite{Beau2026a3}.

  \subsection{Exclusion of abelian realizations}
    \label{ssec:abelian-exclusion}

    \begin{lemma}[Exclusion of the abelian neutral sector]
      \label{lem:abelian-exclusion}
      No abelian group $G$ admits a symmetric generating set $S$ satisfying the neutrality
      condition~\eqref{eq:neutrality} in a faithful irreducible representation of dimension~$2$.
    \end{lemma}

    \begin{proof}
      In a one-dimensional representation, $\chi_{\rho}(s) = \rho(s) \in S^{1}$, so
      $|\rho(s)| = 1$ implies $\chi_{\rho}(s) \neq 0$.
      Hence faithfulness in dimension $1$ is incompatible with~\eqref{eq:neutrality}.
      In dimension $2$, if $G$ is abelian, every irreducible complex representation is
      one-dimensional by Schur's lemma over $\mathbb{C}$; a faithful two-dimensional
      representation of an abelian group decomposes into a direct sum of two
      one-dimensional characters, and the trace of any element $g$ equals
      $\chi_{1}(g) + \chi_{2}(g)$.
      Uniform neutrality across a generating set $S$ implies that the trace vanishes on a
      set generating $G$.
      For an abelian group, characters are multiplicative homomorphisms $\chi_{i} \colon G
      \to \mathbb{C}^{\times}$; faithfulness of $\rho = \chi_{1} \oplus \chi_{2}$ requires
      that for every non-identity $g \in G$ there exists $i$ with $\chi_{i}(g) \neq 1$.
      The neutrality condition forces $\chi_{1}(s) + \chi_{2}(s) = 0$, i.e.\
      $\chi_{2}(s) = -\chi_{1}(s)$, for every generator $s \in S$.
      Since $S$ generates $G$ and characters are multiplicative, this propagates to all of
      $G$: $\chi_{2} = -\chi_{1}$ on $G$.
      But $\chi_{2}$ must be a group homomorphism, so $\chi_{2}(gh) = \chi_{2}(g)\chi_{2}(h)$,
      while $-\chi_{1}(gh) = -\chi_{1}(g)\chi_{1}(h)$.
      These are equal only if $(-1)^{2} = 1$ holds globally, i.e.\ the sign factor $-1$
      introduced by neutrality must itself be multiplicative.
      But a global sign $\chi_{2} = -\chi_{1}$ with $|\chi_{1}(g)| = 1$ for all $g$
      is incompatible with multiplicativity unless $\chi_{1}$ is trivial, contradicting
      faithfulness.
      Therefore no abelian group admits a faithful neutral generating set in dimension~$2$.
    \end{proof}

    \begin{remark}[Scope of the exclusion]
      \label{rem:exclusion-scope}
      Lemma~\ref{lem:abelian-exclusion} excludes all abelian candidates.
      The remaining candidates are finite non-abelian groups admitting a faithful
      irreducible two-dimensional representation.
      By the Klein--Schur classification recalled in~\cite{Beau2026a3}, these are
      exactly the finite subgroups of $\SU(2)$: the dicyclic groups
      $\mathrm{Dic}_{n}$ ($n \geq 2$), the binary tetrahedral group $2T$, the binary
      octahedral group $2O$, and the binary icosahedral group $2I$.
      The binary tetrahedral group $2T$ is further structurally excluded by
      SpectralGramRigidity~\cite{Beau2026a3} (its full neutral sector generates
      only $Q_{8}$, not $2T$).
      The admissible candidates are therefore $Q_{8}$,
      $\mathrm{Dic}_{n}$ ($n \geq 3$), $2O$, and $2I$.
    \end{remark}

  \subsection{Exclusion of octonion-type realizations within the Weil framework}
    \label{ssec:octonion-exclusion}

    \begin{lemma}[Octonion incompatibility with the Weil realisation]
      \label{lem:octonion-exclusion}
      The octonion algebra $\mathbb{O}$ is incompatible with the class of unitary
      representations used by the O-series: the Weil representation on $\Heis$.
    \end{lemma}

    \begin{proof}
      The Weil representation $\rho_{c}$ on $\Heis$ is defined via the cocycle rule
      \[
        \rho_{c}(g_{1} g_{2}) = \phi(g_{1}, g_{2})\, \rho_{c}(g_{1})\, \rho_{c}(g_{2}),
      \]
      where $\phi$ is a $\mathrm{U}(1)$-valued two-cocycle.
      Since $\rho_{c}$ acts on a finite-dimensional complex Hilbert space $V$ via linear
      operators, its representation matrices are elements of $\mathrm{End}(V)$, which is
      an associative algebra by definition of operator composition.
      This construction requires that the algebra of representation matrices be
      \emph{associative}: the cocycle identity forces
      $(\rho_{c}(g_{1})\, \rho_{c}(g_{2}))\, \rho_{c}(g_{3})
      = \rho_{c}(g_{1})\,(\rho_{c}(g_{2})\, \rho_{c}(g_{3}))$.
      The octonion algebra $\mathbb{O}$ is non-associative: there exist
      $x, y, z \in \mathbb{O}$ with $(xy)z \neq x(yz)$.
      A representation taking values in $\mathbb{O}$ cannot satisfy the cocycle identity
      for all triples $(g_{1}, g_{2}, g_{3})$.
      Hence no Weil-type representation on $\Heis$ can be realized over $\mathbb{O}$.
    \end{proof}

    \begin{remark}[Scope of the octonion exclusion]
      \label{rem:octonion-scope}
      Lemma~\ref{lem:octonion-exclusion} does not assert that $\mathbb{O}$ is excluded from
      all possible physical frameworks.
      It asserts only that the specific class of Weil representations used throughout the
      O-series --- which requires associativity of the representation algebra via the
      cocycle identity --- is incompatible with an octonionic realization.
      A broader framework admitting non-associative representations is not excluded in
      principle; it lies outside the scope of the present programme.
    \end{remark}

    \begin{corollary}[Residual classification]
      \label{cor:residual}
      Within the class of unitary Weil representations on $\Heis$, the admissible neutral
      sector is realised by finite subgroups of $\SU(2)$ from the set
      $\{Q_{8},\, \mathrm{Dic}_{n}\ (n \geq 3),\, 2O,\, 2I\}$.
      The underlying real associative normed division algebra is necessarily $\mathbb{H}$.
    \end{corollary}

    \begin{proof}
      By Lemma~\ref{lem:abelian-exclusion} and Remark~\ref{rem:exclusion-scope}, the
      admissible candidates are finite non-abelian subgroups of $\SU(2)$.
      By Lemma~\ref{lem:octonion-exclusion}, the algebra of representation matrices must
      be associative.
      The Hurwitz theorem~\cite{Hurwitz1898} classifies real associative normed division
      algebras as $\mathbb{R}$, $\mathbb{C}$, and $\mathbb{H}$.
      Since $\mathbb{R}$ is one-dimensional (too small for a two-dimensional
      representation) and $\mathbb{C}$ is commutative (excluded by
      Lemma~\ref{lem:abelian-exclusion} at the Lie algebra level), the minimal associative
      non-commutative structure is $\mathbb{H}$.
    \end{proof}

  \section{Quaternionic Minimality and Three Stable Directions}
  \label{sec:quaternionic}

  \subsection{The Hurwitz-to-$\mathfrak{su}(2)$ bridge (imported result)}
    \label{ssec:hurwitz-bridge}

    The following result is Proposition~7.2 of SpectralCapacity~\cite{Beau2026a2},
    reproduced here for self-containedness.

    \begin{proposition}[Hurwitz-to-$\su(2)$ bridge, {\cite{Beau2026a2}}]
      \label{prop:hurwitz-bridge}
      Let $\rho$ be an irreducible two-dimensional complex representation of a finite group
      $G$, and suppose there exist three \emph{linearly independent}, mutually non-commuting
      generators $s_{1}, s_{2}, s_{3} \in G$ (and their inverses) such that
      $\chi_{\rho}(s_{i}) = 0$ for $i = 1, 2, 3$.
      Then the matrices $\rho(s_{i})$ are traceless elements of $\mathrm{SL}(2, \mathbb{C})$,
      and the algebra they generate contains a copy of $\su(2)$ as its real form.
    \end{proposition}

    The proof in~\cite{Beau2026a2} proceeds as follows.
    Each $\rho(s_{i}) \in \SU(2)$ is traceless, so $\rho(s_{i})^{2} = -I$.
    Three mutually non-commuting traceless matrices in $\mathrm{SL}(2, \mathbb{C})$
    satisfying the quaternion relations (which follow from their independence and the
    structure of $\mathrm{SL}(2, \mathbb{C})$) span the Lie algebra $\su(2) \cong
  \mathfrak{sp}(1)$ over $\mathbb{R}$.
    This is the unique associative real structure in real dimension~$3$, by the Hurwitz
    classification~\cite{Hurwitz1898}: among $\mathbb{R}$, $\mathbb{C}$, $\mathbb{H}$,
    $\mathbb{O}$, the last case is non-associative.

  \subsection{Neutral directions as axes of $\ImH$}
    \label{ssec:neutral-as-axes}

    This subsection mediates the identification between ``stable admissible directions''
    and ``generators of $\su(2)$'', a step that must be explicit to avoid conflating
    algebraic and geometric senses of the word ``direction''.

    By equation~\eqref{eq:pauli-axis}, each neutral generator $s \in S$ determines an
    axis $\pm \vec{u}_{s} \in \mathbb{RP}^{2}$.
    The pairwise character geometry of~\cite{Beau2026a3} encodes these axes
    in the Gram matrix
    \begin{equation}
      \label{eq:gram}
      G_{st} = \vec{u}_{s} \cdot \vec{u}_{t} = -\tfrac{1}{2}\,\chi_{\rho}(st),
    \end{equation}
    which is entirely determined by representation-theoretic data.
    The rigidity of~\eqref{eq:gram} under admissible deformations means that the
    relative angles between axes cannot be continuously varied without violating the
    neutrality condition~\eqref{eq:neutrality}: admissible configurations form discrete
    families rather than continuous manifolds.

    A \emph{stable admissible direction} is therefore an axis $\pm \vec{u}_{s}$ that
    belongs to an admissible neutral generating set in the sense of the programme:
    it is a traceless element of $\SU(2)$ compatible with the Born--Infeld capacity
    bound.
    By~\eqref{eq:pauli-axis}, each such axis corresponds to an element of
    $\ImH \cong \mathbb{R}^{3}$ via $\vec{u}_{s} \mapsto u_{s}^{x}\,\mathbf{i}
  + u_{s}^{y}\,\mathbf{j} + u_{s}^{z}\,\mathbf{k}$.
    The number of independent stable admissible directions is thus the rank of the
    neutral traceless subspace over $\mathbb{R}$, which equals
    $\dim_{\mathbb{R}} \ImH = 3$.

  \subsection{Main theorem}
    \label{ssec:main-theorem}

    \begin{theorem}[Three stable directions from quaternionic minimality]
      \label{thm:three-directions}
      Let $\Pi \colon \chi \to \mathcal{O}$ be a non-injective projection compatible with
      the Born--Infeld constraint, admitting a Weil realisation on $\Heis$.
      The number of independent stable admissible directions --- i.e.\ the rank of the
      admissible neutral traceless subspace at the fibre level --- is exactly~$3$.
      Consequently, $\Sigma_{c}(n_{3}) = 3$.
    \end{theorem}

    \begin{proof}
      By O18~\cite{Beau2026a22}, the minimal fibre of $\Pi$ is the involution $c \leftrightarrow
      q - c$ (Theorem~3.3 of O18), derived from the evenness of the Born--Infeld action.
      By Corollary~\ref{cor:residual}, the admissible neutral sector is realised within
      $\mathbb{H}$: the algebra of Weil representation matrices is associative and
      non-commutative, with $\mathbb{H}$ as the minimal such structure.
      By Proposition~\ref{prop:hurwitz-bridge} (Hurwitz-to-$\su(2)$ bridge), three mutually
      non-commuting neutral generators in a two-dimensional irreducible representation span
      a copy of $\su(2)$ as their real form.
      By Section~\ref{ssec:neutral-as-axes}, the independent stable admissible directions
      correspond bijectively to independent axes in $\ImH$.
      Since $\dim_{\mathbb{R}} \ImH = 3$ and this subspace is the maximal subspace of
      $\mathbb{H}$ orthogonal to $\mathbb{R} \cdot \mathbf{1}$ and closed under the
      commutator, the rank of the admissible neutral traceless subspace is exactly~$3$.
      Each independent axis of $\ImH$ contributes one independent neutral mode to the
      admissible Gram--Schmidt span; saturation occurs when all three independent axes have
      been realised by the projection.
      The threshold condition $\Sigma_{c}(n_{3}) = 3$ in O21~\cite{Beau2026a25} records
      this rank as an observable: the capacity count $\Sigma_{c}$ reaches the value
      $3$ precisely at the shell $n_{3}$ where all three independent neutral directions
      have been saturated by the projection.
    \end{proof}

    \begin{remark}[Conditionality inherited from O18]
      \label{rem:conditionality}
      Theorem~\ref{thm:three-directions} inherits the conditionality of O18 Theorem~2.5:
      the derivation assumes that Born--Infeld parity is the only global effective symmetry
      of $S_{\mathrm{BI}}[\chi]$ compatible with BI-admissible observables.
      If the effective action admits additional symmetries (topological, gauge, or other),
      the fibre could be larger, and additional neutral directions could be present,
      modifying the count.
      A general classification of fibres in enriched frameworks is deferred to a subsequent
      paper following O21, as noted in O18~§4.3.
    \end{remark}

    \begin{remark}[The ``$3$'' is not the dimension of $\mathbb{H}$]
      \label{rem:not-four}
      The integer $3$ derived in Theorem~\ref{thm:three-directions} is $\dim_{\mathbb{R}}
      \ImH$, not $\dim_{\mathbb{R}} \mathbb{H} = 4$.
      The full quaternion algebra contributes one additional real dimension ($\mathbb{R}
      \cdot \mathbf{1}$), which is commutative and plays no role in the neutral traceless
      sector.
      The Born--Infeld neutrality condition~\eqref{eq:neutrality} projects out this
      commutative direction, leaving exactly three independent axes.
    \end{remark}

    \begin{remark}[Causal direction: $\ImH$ precedes $\su(2)$]
      \label{rem:causal-order}
      The coincidence $\dim \su(2) = 3$ is not the origin of the result but a consequence
      of the quaternionic realisation.
      The logical order is: Born--Infeld admissibility forces $\mathbb{H}$-minimality;
      the imaginary subspace $\ImH \cong \mathbb{R}^{3}$ is then the unique neutral
      traceless subspace; and $\su(2) \cong \ImH$ as real Lie algebras follows from the
      bracket structure of $\ImH$.
      Appealing directly to $\dim \su(2) = 3$ without this chain would be circular: it
      would assume the $\mathbb{H}$-structure rather than deriving it.
    \end{remark}

  \section{Spectral Realisation and Consistency with the Programme}
  \label{sec:spectral}

  \subsection{$Q_{8}$ as the minimal isotropic prototype}
    \label{ssec:q8}

    The quaternion group $Q_{8}$ is the smallest finite subgroup of $\SU(2)$ admitting a
    neutral generating set, and SpectralGramRigidity~\cite{Beau2026a3} identifies
    it as the unique admissible group with an orthogonal Gram pattern
    ($G_{st} = 0$ for all $t \neq s^{\pm 1}$).
    Its six traceless elements $\{\pm\mathbf{i}, \pm\mathbf{j}, \pm\mathbf{k}\}$
    pair into three conjugate pairs, each determining one of the three coordinate axes
    of $\ImH$.

    $Q_{8}$ therefore provides a discrete, exact prototype of the abstract derivation:
    it realises precisely three independent neutral axes, isotropically distributed, with
    the second-moment tensor $M \propto I$~\cite{Beau2026a3}.
    The three neutral axis directions of $Q_{8}$ are the finite-group realisation of the
    three generators of $\su(2)$, confirming the identification of
    Section~\ref{ssec:neutral-as-axes} in the simplest admissible case.

  \subsection{Three-level ADE stratigraphic structure}
    \label{ssec:ade-three-levels}

    For the ADE binary Cayley graphs physically relevant to the Cosmochrony
    programme --- in particular $2I$ with order-$4$ and order-$5$ generators --- the
    Cayley eigenvalue formula gives, for each irreducible representation $\rho$ of $G$,
    \[
      \lambda_{\rho} = |S| - \frac{1}{d_{\rho}} \sum_{s \in S} \chi_{\rho}(s).
    \]
    The character sum $\sum_{s \in S} \chi_{\rho}(s) / d_{\rho}$ takes exactly three
    distinct values over the non-trivial irreducible representations of $G$, corresponding
    to three distinct eigenvalue classes $\lambda_{1} < \lambda_{2} < \lambda_{3}$.
    This produces exactly three non-zero Cayley eigenvalue classes, and hence three
    stratigraphic levels~\cite{Beau2026a4}.

    Theorem~\ref{thm:three-directions} provides the structural explanation for this
    observation: the three spectral levels are the spectral realisation of the three
    independent axes of $\ImH$.
    The algebraic argument derives the ``$3$'' as necessary; the ADE spectral structure
    shows how this ``$3$'' is realised concretely in the relevant substrate graphs.
    The two arguments are complementary and non-circular: the algebra explains why three
    is necessary; the ADE confirms how three is achieved.

  \subsection{Consistency with $\Sigma_{c}(n_{3}) = 3$ in O21}
    \label{ssec:o21-consistency}

    O21~\cite{Beau2026a25} defined $n_{3}$ as the BFS shell at which $\Sigma_{c}$ reaches
    $3$, and O22 showed that projection locking forces saturation onto this shell.
    Theorem~\ref{thm:three-directions} now closes the logical chain: the value $3$ is the
    rank of $\ImH$ as the admissible neutral traceless subspace of the Born--Infeld
    fibre.
    The full chain of derivation is:
    \begin{equation}
      \label{eq:chain}
      \text{BI parity}
      \;\Rightarrow\; \text{fibre involution } c \leftrightarrow q-c
      \;\Rightarrow\; \text{non-abelian admissible support}
      \;\Rightarrow\; \mathbb{H}\text{-minimality}
      \;\Rightarrow\; \dim \ImH = 3
      \;\Rightarrow\; \Sigma_{c}(n_{3}) = 3.
    \end{equation}

  \section{Discussion}
  \label{sec:discussion}

  \subsection{What O23 derives and what it does not}
    \label{ssec:scope-discussion}

    O23 derives the \emph{value} $3$ in the threshold condition $\Sigma_{c}(n_{3}) = 3$
    as a structural consequence of Born--Infeld fibre admissibility and quaternionic
    minimality.
    It does not determine the \emph{location} $n_{3}$ as a function of the pipeline
    parameters $(c_{\mathrm{BI}}, \lambda_{n}, \text{graph geometry})$.
    That question requires the extended numerical campaign described in O22~§6.2 and
    remains open.

    The position of O23 in the O-series is therefore a structural closure of the
    ``why $3$?'' question, complementary to O22's structural closure of the
    ``why a shell?'' question.
    Together, O22 and O23 complete the derivation of the observable $n_{3}$ at the
    structural level, leaving the quantitative determination of its numerical value as
    the remaining open task.

  \subsection{Conditional status and possible extensions}
    \label{ssec:conditional}

    The main result carries the conditionality of O18 Remark~2.6 (recalled in
    Remark~\ref{rem:conditionality} above).
    If the Born--Infeld effective action admits additional symmetries beyond parity ---
    arising, for instance, from topological constraints identified in the Topological
    Invariants paper~\cite{Beau2026k} or from gauge structure in an enriched framework
    --- the fibre could be larger and the neutral traceless subspace could have a higher
    rank.
    Determining whether such extensions are realised in the Cosmochrony framework, and
    if so what rank they imply, is deferred to the general fibre classification programme
    announced in O18~§4.3.

  \section{Conclusion}
  \label{sec:conclusion}

  We have derived the threshold condition $\Sigma_{c}(n_{3}) = 3$ of
  O21~\cite{Beau2026a25} from Born--Infeld fibre admissibility and quaternionic
  minimality.
  The central result is Theorem~\ref{thm:three-directions}: within the class of Weil
  representations on $\Heis$ compatible with Born--Infeld parity, the admissible
  neutral traceless subspace is isomorphic to $\ImH$, and its real dimension is
  exactly~$3$.

  The derivation rests on three pillars: the exclusion of abelian realizations
  (Lemma~\ref{lem:abelian-exclusion}); the incompatibility of octonion-type
  realizations with the Weil cocycle structure (Lemma~\ref{lem:octonion-exclusion});
  and the Hurwitz-to-$\su(2)$ bridge (Proposition~\ref{prop:hurwitz-bridge},
  imported from SpectralCapacity~\cite{Beau2026a2}).

  The result is conditional on the absence of additional effective symmetries beyond
  Born--Infeld parity, in accordance with O18~\cite{Beau2026a22}.
  Subject to this condition, it closes the last structural open problem of O21: the
  ``$3$'' in $\Sigma_{c}(n_{3}) = 3$ is not a phenomenological input but
  \emph{the real dimension of the imaginary quaternions as the minimal admissible
  non-commutative associative neutral structure forced by Born--Infeld parity}.
  The emergence of three stable directions is therefore not geometric but algebraic
  in origin.

  \bigskip
  \noindent
  \textbf{Acknowledgements.}
  This paper is part of the Spectral Admissibility Sub-programme of the Cosmochrony
  framework~\cite{Cosmochrony}.

  \appendix

  \section{Why Not 2? Why Not 7?}
  \label{app:not-2-not-7}

  The exclusion argument of Section~\ref{sec:exclusion} can be summarised concisely by
  examining what happens at the other dimensions allowed by the Hurwitz classification.

  \paragraph{Why not 2 (complex numbers $\mathbb{C}$)?}
    The algebra $\mathbb{C}$ is a two-dimensional real division algebra, but it is
    \emph{commutative}.
    A representation taking values in $\mathbb{C}^{\times}$ is one-dimensional over
    $\mathbb{C}$, i.e.\ abelian, and is excluded by Lemma~\ref{lem:abelian-exclusion}.
    In the language of Section~\ref{ssec:neutral-as-axes}, the imaginary subspace
    $\mathrm{Im}\,\mathbb{C} = \mathrm{i}\mathbb{R}$ has real dimension~$1$: only one
    neutral axis is available, which is insufficient to carry three independent stable
    directions.

  \paragraph{Why not 7 (imaginary octonions)?}
    The imaginary subspace $\mathrm{Im}\,\mathbb{O}$ has real dimension~$7$, which would
    naively suggest $7$ stable directions.
    However, $\mathbb{O}$ is non-associative, and Lemma~\ref{lem:octonion-exclusion}
    shows that this is incompatible with the Weil cocycle structure: no representation
    of $\Heis$ can take values in $\mathbb{O}$ while satisfying the cocycle identity.
    The $7$-dimensional imaginary space is therefore inaccessible within the
    Born--Infeld Weil framework.

  \paragraph{Summary.}
    The Hurwitz classification $\{\mathbb{R}, \mathbb{C}, \mathbb{H}, \mathbb{O}\}$ has
    exactly one entry that is associative and non-commutative: $\mathbb{H}$.
    Its imaginary subspace has dimension~$3$.
    The Born--Infeld neutrality condition selects this unique entry, and the resulting
    count $\dim_{\mathbb{R}} \ImH = 3$ is forced.

    \clearpage
    \bibliographystyle{plain}
    \bibliography{cosmochrony-bibliography,external-refs}

\end{document}
